---
title: "Homework 1"
author: "Lauren Littig (LL3969)"
output: html_document
date: "2026-09-20"
---

# Problem 1

```{r}
library(tidyverse)
data("penguins", package = "palmerpenguins")
```

The `penguins` dataset contains measurements and characteristics of penguins from three species (Adelie, Chinstrap, and Gentoo) observed on the islands of Biscoe, Dream, and Torgersen. Important variables include `species`, `island`, `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_g`, `sex`, and `year`. 

The dataset contains `r nrow(penguins)` observations and `r ncol(penguins)` variables. The mean flipper length is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)`

```{r}
ggp_penguins = 
  ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm)) + 
  geom_point(aes(color = species))

ggsave("ggp_penguins.pdf", ggp_penguins)
```

# Problem 2

```{r}
samp_df = 
  tibble(
    random_samp = rnorm(10),
    vec_logical = random_samp > 0,
    vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    vec_factor = factor(
      c("a", "b", "c", "a", "b", "c", "a", "b", "c", "a")
    )
  )

mean(pull(samp_df, random_samp))
mean(pull(samp_df, vec_logical))
mean(pull(samp_df, vec_char))
mean(pull(samp_df, vec_factor))

```

I cannot take the mean of `vec_char` because character strings cannot be averages, nor can I take the mean of `vec_factor` because factors are categorical variables.

```{r results = "hide"}
as.numeric(pull(samp_df, vec_logical))
as.numeric(pull(samp_df, vec_char))
as.numeric(pull(samp_df, vec_factor))
```

Applying `as.numeric()` to the logical variable converts `TRUE` and `FALSE` to 1 and 0, respectively. Applying `as.numeric()` to the character variable produces `NA` values because the character strings cannot be interpreted as numbers. Applying `as.numeric()` to the factor variable returns the underlying integer codes assigned to its factor levels rather than the original labels. This helps explain why taking the mean of the logical variable works, while taking the mean of the character and factor variables does not directly produce a meaningful result.


